\documentclass[a4paper, 11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{proof}
\usepackage{color}
\usepackage{url}

\title{Informal system description}
\author{Giselle Reis}

\begin{document}
\maketitle

A quick diagram of what is implemented so far (and available at
\url{http://www.logic.at/people/giselle/tatu}) is the following:

\vspace{1cm}

\begin{tabular}{|p{1.7cm}|}
\hline
Sequent calculus\\
\hline
\end{tabular}
$\overset{(a)}{\Longrightarrow}$
\begin{tabular}{|p{2.3cm}|}
\hline
SELLF specification\\
\hline
\end{tabular}
$\Longrightarrow$
\begin{tabular}{|p{1.5cm}|}
\hline
\textsf{\textbf{TATU}}\\
\hline
\end{tabular}
$
\left\{
  \begin{array}{l}
  \text{\scriptsize{Principal cut (1)}}\\
  \text{\scriptsize{Atomic cut elimination (2)}}\\
  \text{\scriptsize{Cut-coherence (3)}}\\
  \text{\scriptsize{Initial coherence (4)}}\\
  \end{array}
\right.
$

\vspace{1cm}

Quick explanation of each functionality:

\begin{enumerate}
  \item \textbf{Principal cut}: Checks if cuts and rules can be permuted in such
  a way that cuts become principals. This is done by checking which subexponential
  indices are used in the specification of the rules (cut and introduction
  rules).
  
  \item \textbf{Atomic cut elimination}: Checks whether atomic cuts can be
  eliminated. This is also done by checking which subexponentials indices are
  used in the specification of the rules (cut and introduction rules).
  
  \item \textbf{Cut-coherence}: Checks whether cuts on some formula with
  $\diamond$ as its top-level connective can be replaced by simpler cuts. This
  is done by checking the duality of the specifications for the introduction
  rules of $\diamond$. Let $B_r$ and $B_l$ be the specification of the right and
  left introduction rules for $\diamond$, respectively. Then we must prove: $\vdash Cut,
  B_r^{\perp} \bindnasrepma B_l^{\perp}$.

  \item \textbf{Initial coherence}: Checks wheter initial rules on some formula
  with $\diamond$ as its top-level connective can be replaced by simpler initial
  rules. Again, this is done by checking some kind of duality of the formulas.
  Let $B_r$ and $B_l$ be the specification of the right and left introduction
  rules for $\diamond$, respectively. Then we must prove $\vdash Init,
  ?^{\infty} B_r \bindnasrepma ?^{\infty} B_l$.
\end{enumerate}

For numbers 3 and 4, a bounded proof seach procedure was implemented (the bound
is set to 4 by default).

In order to implement the proof search procedure, the SELLF system was adapted with
input and output contexts, following the same ideas of ``Efficient Resource
Management for Linear Logic Proof Search'', by Cervesato, Hodas and Pfenning.
This is the system in Figure \ref{figure:sellf}. These contexts are used to
avoid the non-determinism of splitting the contexts in the $\otimes$-rule.

Axioms set the output context. This should be empty when the procedure returns
to the end sequent. In-contexts are set bottom up and out-contexts are set
top down (when returned from the recursive call).

{\color{red}Explain initial rule, top, one and bang and tensor.}

\input{sellfimplemented}

\section*{Future work}

\begin{itemize}
  \item Part $(a)$ of the diagram is done completely by hand. Currently Vivek
  and Ramyaa are developing an \textbf{algorithm to automatize the specification
  process}.

  \item A nice feature would be to \textbf{print a latex code of the system specified in
  SELLF} (would be nice to validate the specification). This can be done using
  the macro-rules of each specification formula and translating them to the object
  logic rule. The generation of macro-rules was implemented at some point, and
  the system actually printed this macro-rule in latex. I would have to check
  again if this code still works and to translate the derivation into a single
  rule.
  
  \item Considering that the specification has a high-level of adequacy, the
  \textbf{proof search in SELLF can be used to do proof search on the object
  logic}. This
  is relatively straightforward to implement and I want to do this for Anna's
  system for paraconsistent logic. We need to uncomment the querying part and
  debug it.

  \item Some time ago, I and Vivek had a project that \textbf{checked automatically
  whether rules of the object logic permute} by making an exhaustive search and
  using answer set programming. The code still exists but there are some bugs
  that we noticed afterwards. This needs to be fixed, but I am afraid that it
  will involve changing the data structure.

  \item Given that there is a way of checking automatically which rules permute
  over which, we could build a permutation graph and try to \textbf{deduce a focused
  proof system for the object logic} from it. This is just an idea and the
  theory behind it should be developed first.

  \item Given that our system proves a lot of things, we think about having
  \textbf{proof objects}. But this is a long term goal, and something that would
  fit Dale Miller's ProofCert project.
\end{itemize}

\end{document}
