\documentclass{llncs}
\usepackage{latexsym}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{proof}
\usepackage{comment}
\usepackage{mathbbol}

\DeclareMathAlphabet{\mathsl}{OT1}{cmr}{m}{sl}
\newcommand\tsl[1]{\hbox{\ensuremath{\mathsl{#1}}}}

\newcommand{\sellf}{\hbox{SELLF}}
\newcommand\zero{0}
\newcommand\one{1}
\newcommand\bottom{\perp}
\newcommand\bang{\mathop{!}}
\newcommand\quest{\mathord{?}}
\newcommand\limp{\mathbin{-\hspace{-0.70mm}\circ}}
\newcommand\tensor\otimes
\newcommand\bla{\mathrel{\mbox{$\circ\!-$}}}
\newcommand\lftall{\forall_{l}}
\newcommand\lexists{\exists_{l}}
\newcommand\lolli{\multimap}
\newcommand\plus{\oplus}
\newcommand\lpar{\mathrel{\bindnasrepma}}
\newcommand\bigpar{\mathrel{\bindnasrepma}}
\newcommand{\with}{\mathbin{\&}}

\newcommand{\nbang}[1]{\hbox{$\bang^\mathsl{#1}$}}
\newcommand{\nquest}[1]{\hbox{$\quest^\mathsl{#1}$}}

\newcommand{\unb}{\tsl{unb}}

\newcommand{\fail}{\hbox{\tsl{fail}}}
\newcommand{\mctx}[2]{\hbox{\tsl{mctx}(\ensuremath{#1},\ensuremath{#2})}}
\newcommand{\elin}[2]{\hbox{\tsl{elin}(\ensuremath{#1},\ensuremath{#2})}}
\newcommand{\emp}[1]{\hbox{\tsl{emp}(\ensuremath{#1})}}
\newcommand{\eqf}[2]{\hbox{\tsl{eqf}(\ensuremath{#1},\ensuremath{#2})}}
\newcommand{\addform}[3]{\hbox{\tsl{addform}(\ensuremath{#1},\ensuremath{#2},\ensuremath{#3})}}
\newcommand{\eqctx}[2]{\hbox{$
\tsl{eqctx}$(\ensuremath{#1},\ensuremath{#2})}}
\newcommand{\union}
[3]{\hbox{\tsl{union}(\ensuremath{#1},\ensuremath{#2},\ensuremath{#3})}}
\newcommand{\In}
[2]{\hbox{\tsl{in}(\ensuremath{#1},\ensuremath{#2})}}
\newcommand{\form}
[1]{\hbox{\tsl{form}(\ensuremath{#1})}}
\newcommand{\Equ}
[2]{\hbox{\tsl{Equ}(\ensuremath{#1},\ensuremath{#2})}}


\newcommand{\mr}[3]{\tsl{mr}(#1,#2,#3)}
\newcommand{\der}[7]{\tsl{mr^2}(#1,#2,#3,#4,#5,#6,#7)}

\newcommand{\Oscr}{\mathcal{O}}
\newcommand{\Cscr}{\mathcal{C}}
\newcommand{\Lscr}{\mathcal{L}}
\newcommand{\Sscr}{\mathcal{S}}
\newcommand{\Tscr}{\mathcal{T}}
\newcommand{\Wscr}{\mathcal{W}}

% For generic contexts
\newcommand{\genK}{\mathcal{K}_{g}}
\newcommand{\genG}[2]{\Gamma_{#1}^{#2}}

\newcommand{\eg}{{\em e.g.}}
\newcommand{\ie}{{\em i.e.}}
\newcommand{\etal}{\emph{et al.}}

\newcommand{\alg}[1]{\mathbb{\Lbrack} #1 \mathbb{\Rbrack}}

% Evironments for comments
\newcommand{\Endmark}{$\diamond$}
\newenvironment{giselle}{\begin{trivlist}\item[]{\bf Giselle:\ }}%
                       {{\hfill \Endmark}\end{trivlist}}

\newenvironment{vivek}{\begin{trivlist}\item[]{\bf Vivek:\ }}%
                       {{\hfill \Endmark}\end{trivlist}}

%opening
\title{Automating the Proof of Permutation Lemmas}
\author{Vivek Nigam \texttt{vivek.nigam at ifi.lmu.de}\\
  Giselle Reis \texttt{giselle at logic.at}}

\institute{Ludwig-Maximilians-Universit{\"a}t, Munich, Germany\\
Technische Universit{\"a}t Wien, Vienna, Austria}


\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}

\input{intro}
\section{Linear Logic}

Focusing;
Subexponentials.

\section{Encoding Proof Systems}

Linear logic with subexponentials as a logical framework for proof systems.

\section{Macro-rules}

\input{macro}

\subsection{Representing Sequents: generic contexts}
\label{sec:generic_contexts}

\input{gen_contexts}

\begin{comment}
\begin{giselle}
Maybe we don't need to make a distinction between the unbounded context or the
others... since it is declared in the specification just like the other
subexponentials. Do we?
\end{giselle}

We assume that the names of the contexts in the object-logic are specified
as subexponential names and that the meta-logic has an unbounded
subexponential, referred to as \unb, that may only contain the clauses specifying 
the inference rules of the object-logic.

A side-formula context is represented as a function from the set of
subexponential 
indexes to a set of \sellf\ formulas and generic context names, such as 
$\Gamma$. 
\end{comment}

\subsection{Constraints}
\label{sec:constraints}

\input{constraints}

\subsection{Macro-rules}
\label{sec:macro_def}

\input{macro_def}

\begin{comment}
Then, the set of macro-rules,
$\mr{F}{\Cscr}{\Oscr}{\Tscr}{\Sscr}$, that can introduce a formula, $F$,
given a root side-formula context $\Sscr$, is represented by a list of
closed leaves
$\Cscr$, a list of open leaves $\Oscr$, and a set of sets of constraints,
$\Tscr$.
\end{comment}

\begin{comment}
Given a formula, $F$, and a root side-formula context, $\Sscr$, we
macro-rules of a formula inductively. Our procedure will have an invariant
that the unbounded contexts will have exactly one generic context. This
does not mean, however, that an unbounded context might not have more than 
a formula. 

There are two base cases: The first base case is when a derivation
finishes with an initial rule. In this case $F = A^\bot$. The set of closed
and open leaves are, respectively,  $\Cscr = \{S\}$ and $\Oscr
= \emptyset$. We construct the set of constraints as follows. Assume that
$u_1, \ldots, u_n$ are the unbounded
subexponentials and $\Lscr = l_1, \ldots, l_m$ are the linear
subexponentials. In order for the initial rule to be applicable $A$ must be
present in some context of the side-formula context $\Sscr$. There are two 
cases, $A$ appears in an unbounded context or $A$ is the only element in 
a linear context. 

If $A$ appears in an unbounded context $\tsl{u}_j$, then 
all the linear contexts must contain only generic contexts, $\Gamma_{l_i}$.
% G: Why generic contexts? Can't we just say that all linear context must be
% empty? They are always generic in some sense...
% VN: Here what I meant is that no formulas can appear in linear contexts. 
% Remember that we have two types of objects that can appear in contexts:
% generic contexts and formulas. 
We construct the constraints specifying that the linear contexts are all
empty as follows: $\Tscr_l = \{\emp{\Gamma_i} \mid i \in \Lscr \}$.
If any linear context contains a formula, then the macro-rule is not valid.
% G: Do we really have to include this fail by ourselves? Won't smodels realize
% it's unsatisfiable if there is a constraint elin(X, \Gamma_i)?
% VN: There are two ways of handling this. One is transforming Gamma, F
% into a new generic context Delta, and then specifying that emp(Delta)
% or we do this by hand by using the fail constraint.
% GR: Do you mean that if we have the following constraints co-existing: elin(X,
% \Gamma_i) and emp(\Gamma_i) it will not fail by itself?
In this case,  $\Tscr_l = \{ \fail \}$, that is, it is macro-rule that 
is not satisfiable.
Moreover, if the context for $\tsl{u}$ is $\Gamma_u, F_1, \ldots, F_{k_u}$,
then there are $k_u+1$ possible macro-rules: either $A \in \Gamma_{u_j}$ or
$A = F_i$, for $1 \leq i \leq k_u$. These possible macro-rules are again
specified by using
constraints:
$\Tscr_{u_j}^{\Gamma_u} = \{\mctx{F}{\Gamma_u}\}$, and
$\Tscr_{u_j}^i = \{\eqf{F}{F_i}\}$ for $1 \leq i \leq k$.

\[\Tscr_u = \bigcup_{1\leq j\leq n} \left\{\{\Tscr_{u_j}^{\Gamma_{u_j}}
\cup \Tscr_l \} \cup \bigcup_{1 \leq i \leq k_{u_j}}\{\Tscr_{u_j}^i \cup
\Tscr_l\}\right\}.\]

If $A$ is in a linear context, say $l_j$, then $A$ must be the only 
formula in that context. Hence, the context of $l_j$ may contain 
a generic context, $\Gamma_{l_j}$, or a context and a single formula
$F_{l_j}$. 
In the former case, $\Gamma_{l_j}$ contains only one formula $A$, which 
is specified by the constraint
$\elin{F}{\Gamma_{l_j}}$, while in the latter case, $\Gamma_{l_j} =
\emptyset$ and $A = F_{l_j}$, which is specified by the constraints 
$\{\emp{\Gamma_{l_j}}, \eqf{A}{F_{l_j}}\}$. Moreover, all the other 
linear context must be empty, which is specified by the set of constraints
$\Tscr_{l_j} = \Tscr_l \setminus \{\emp{\Gamma_{l_j}}\}$. If any
context contains a formula, then the macro-rule is not valid. In this
case,  $\Tscr_{l_j} = \{ \fail \}$. 
Hence, we obtain the following set of set of constraints:

\[\Tscr_l = \bigcup_{1\leq j\leq m} \{\{\elin{F}{\Gamma_{l_j}}\} \cup
\Tscr_{l_j}\} \cup \{\{\emp{\Gamma_{l_j}}, \eqf{A}{F_{l_j}}\} \cup
\Tscr_{l_j}\}.\]
Finally the set of constraints is $\Tscr = \Tscr_u \cup \Tscr_l$. 

% G: Should I stop if I encounter a positive formula??
% I thought we were stopping when we got to atomic principal formulas, since
% this means that the whole clause was decomposed...
The
other base cases result at the end of the 
negative phase, in which case the set of closed and open leaves are,
respectively, $\Cscr = \emptyset$ and $\Oscr = \{S\}$. Moreover, there
are no constraints, that is, $\Tscr = \emptyset$.
\end{comment}

\begin{comment}
For the negative rules, there are no constraints generated nor does
one need to change the root side-formula context.  One
simply appends the list of open and closed leaves, whenever needed, \eg, 
for the $\with$ rule. 
\begin{giselle}
Not entirely true... rule ?l generates a constraint.
\end{giselle}
\end{comment} 

\begin{comment}
For the positive rules, each requires some special 
attention. In particular, for the introduction rule for the existential 
quantifier, one simply instantiates the existentially quantified variable 
with a new variable, so that the resulting formula is uniquely identified. 
The bang introduction rule one changes the root side-formula context of
the premise
so that the side condition of the bang rule is satisfied. In the process
some contexts for unbounded contexts are deleted and some new constraints
of the form $\emp{\Gamma_l}$ might be added for linear contexts
$\Gamma_l$. If a linear context that should be empty contains a formula, 
then the constraint \fail\ is added instead.

The case that one needs to take a bit more care is with the $\tensor$
introduction rule. In particular,  we create a pair of new
generic context for each linear 
subexponential. They are split among the premises of the tensor. On the 
other hand, the contexts for the unbounded contexts are copied to the
premises. We illustrate this case with an example.
Assume that there are only three
subexponentials, one unbounded \tsl{u} and two linear \tsl{l_1} and 
\tsl{l_2}:

% G: Do we need to show the formulas explicitly? Why?
\[
 \infer[\tensor]{\vdash \Gamma_u : \Gamma_{l_1}, \Delta_1 :
\Gamma_{l_2}, \Delta_2: \cdot \Downarrow F_1 \tensor F_2}
{
\vdash \Gamma_u : \Gamma_{l_1}' : \Gamma_{l_2}': \cdot \Downarrow F_1
&
\vdash \Gamma_u : \Gamma_{l_1}'' : \Gamma_{l_2}'': \cdot \Downarrow F_2
}
\]
where $\Delta_i$ are sequences of formulas. 
Notice that the unbounded context $\Gamma_u$ is copied among the premises
and the
new linear contexts $\Gamma_{l_1}', \Gamma_{l_1}'', \Gamma_{l_2}',$ and
$\Gamma_{l_2}''$ are split among the premises. Moreover, that the
formulas in $\Delta_i$ do not appear anymore in the premises. 

For the set of open and closed leaves, we proceed as before and
simply append the set of open and closed leaves obtained using the premises
above as root side-formula contexts. 

For the set of constraints, we proceed in two steps. First, we merge the
constraints obtained by the rule's premises as follows: Assume
$\Tscr_1$ and $\Tscr_2$ are the set of
sets of constraints of the premises of the tensor rule. Then the
constraints of the tensor should include the set:
\[
 \Tscr' = \{T_1 \cup T_2 \mid T_1 \in \Tscr_1 \textrm{ and } T_2 \in
\Tscr_2\}
\]
Intuitively, any macro-rule introducing first premise together with
any macro-rule introducing the second premise is a valid macro-rule for 
the tensor. 

However, we still need to express the relation between the linear 
contexts in the conclusion side-formula context and the new linear contexts
in the premise side-formula contexts. In particular, we want to specify
that $\Gamma_{l_i} \cup \Delta_i = \Gamma_{l_i}' \cup \Gamma_{l_i}''$, 
which could be specified by
constraints of the form
$\eqctx{\{\Gamma_{l_i}, \Delta_i\}}{\{\Gamma_{l_i}', \Gamma_{l_i}'' \}}$.
However, the formulas in $\Delta_i$ are not generic contexts and therefore
this constraint is not well-formed. Therefore, we pre-process 
the set $\Gamma_{l_i} \cup \Delta_i$ so to obtain a set containing only 
generic contexts. This is done by the following operation that introduces
inductively new generic contexts:
if $\Delta_i = \emptyset$ then we are done; otherwise let $F \in
\Delta_i$. Then we introduce two new generic contexts $\Gamma_{l_i}^1$ and 
$\Gamma_{l_i}^F$,
such that $\Gamma_{l_i}^1 = \Gamma_{l_i} \cup \Gamma_{l_i}^F$
specified by the
constraint \union{\Gamma_{l_i}}{\Gamma_{l_i}^F}{\Gamma_{l_i}^1} and that 
$\Gamma_{l_i}^F = \{F\}$, specified by the constraint
\elin{F}{\Gamma_{l_i}^F}. We repeat this operation to the set 
$\Gamma_{l_i}^1 \cup \Delta_i\setminus\{F\}$ that contains less formulas. 
At the end we add all the resulting constraints to all the sets of
constraints in the set $\Tscr'$. 
\end{comment}

\input{example}

\subsection{Soundness and Completeness}

\input{proofs_macro}

%\paragraph{Soundness and Completeness}

%\textbf{VN: Perhaps to make things a bit more clearer, we need a notion of 
%$\alpha$-equivalence among sequent contexts.}

%Let $F$ be a \sellf\ formula. Then, every satisfiable instance of the macro-rules generated
%using the algorithm proposed corresponds to a derivation in \sellf\ when
%focusing on $F$.

%\vspace{0.5cm}

% G: Can I say that macro-rules with unsatisfiable constraints corresponds to
% derivations that fail at some point?

% Yes, this is what the soundness and completeness results will imply.

%  VN: Since there is no negation, there is a least model using standard 
%  Horn theory semantics. There are constraints, but those do not
% introduce 
% % models.
% 

% The following lemma is indeed not correct. The assumption that the 
% our constraints are Horn theories is false, due to the specification of 
% union constraints.

% \begin{lemma}
%  Let $\Tscr$ be a set of constraints of the form above. If $\Tscr$ is 
% satisfiable, then it has a model.
% \end{lemma}

% G: It's not a least model anymore. Check this proof.

%To formalize the soundness of our translation, that takes a
%model of a set of constraints, $\Cscr$, the root side-formulas, and the
%list of open leaves and returns the macro-rule using the least model of
%$\Cscr$. 

% TODO: verify this after making the necessary changes on the example.
\begin{comment}
We illustrate using our
previous example how this translation works. Consider the first set of
constraints. 

\[
\Tscr = \left\{\begin{array}{c}
\emp{\Gamma_r'},\emp{\Gamma_l'}, \mctx{A}{\Gamma_u},
\union{\Gamma_r}{\Gamma_F}{\Gamma_r^1},\\
\elin{F}{\Gamma_F}, \union{\Gamma_r'}{\Gamma_r''}{\Gamma_r^1},
\union{\Gamma_l'}{\Gamma_l''}{\Gamma_l}, \emp{\Gamma_l''}
\end{array}\right\} 
\]
These constraints are satisfiable. To determine its least model, we use the
smodels logic program engine~\cite{niemela97lpmnr}, obtaining the following
model:
\[
 \left\{\begin{array}{c}
\In{F}{\Gamma_r^1},\In{F}{\Gamma_F}, \In{F}{\Gamma_r''},
\In{A}{\Gamma_u}\\
\end{array}\right\}  \cup \Tscr
\]
We then use this model to trigger the rewrite rules of Figure
\ref{fig:rewritting} to the 
contexts appearing in the open leaves and the root side-formulas:

\[
 \begin{array}{l@{\quad}l}
  \emp{\Gamma} & \Gamma \rightarrow \emptyset\\
  \mctx{F}{\Gamma} & \Gamma \rightarrow \Gamma, F\\
  \union{\Gamma'}{\Gamma''}{\Gamma} & \Gamma \rightarrow \Gamma',
\Gamma''\\
  \In{F}{\Gamma} & \Gamma \rightarrow \Gamma, F\\
  \top & \emptyset, \Gamma \rightarrow \Gamma
 \end{array}
\]

The last rewrite is always applicable. Notice that for the other
constraints, such as \eqctx{\cdot}{\cdot}, we do not need rewrite
rules. These are used only for specifying the model.

\vspace{0.5cm}

\begin{lemma}
Let $mr(F) = \langle \Cscr, \Oscr, \Tscr \rangle$ be a macro-rule computed as shown
above. Let $M$ be an arbitrary model of a set of constraints in $\Tscr$
that is satisfiable.
% GR: which translation? the rewritting?
Then the macro-rule obtained by applying the translation shown above using 
$\Oscr, \Sscr$ and $M$ is derivable in \sellf\ by focusing on the
formula $F$. 
\end{lemma}

\begin{proof}
The proof follows by induction on the size of $F$.
The model of $\Tscr'$ is then used as an oracle on the non-deterministic
choices, such as how to do the splitting of contexts.

The first base case is when $F$ is a positive atomic formula. Then the set
of constraints
obtained from the macro-rule algorithm above is of the following form:
\[
 \In{F}{\Gamma} \quad \textrm{and} \quad \emp{\Gamma'}
\]
where $\Gamma$ is a context in the root side-formulas. All of these
constraints are satisfiable. Moreover, it returns an empty set of open leaves.
Then, the translation from (the) of a constraint model to a macro-rule
replaces a generic context $\Gamma$ appearing in an \In{F}{\Gamma}
constraint by $\Gamma, F$ and a generic context $\Gamma'$ appearing 
in a \emp{\Gamma'} by $\emptyset$. The
resulting 
macro-rule corresponds exactly to the a macro-rule when an initial rule is
applied. The second base case, which is when 
$F$ is a negative atomic formula, is straightforward, as no constraints 
are generated.

For the inductive cases, the most interesting case is when $F = F_1
\tensor F_2$. Assume that $\mr{F_1}{\Cscr_1}{\Oscr_1}{\Tscr_1}{\Sscr_1}$
and $\mr{F_2}{\Cscr_2}{\Oscr_2}{\Tscr_2}{\Sscr_2}$ are given, which encode
the macro-rules computed by introducing the subformulas of $F$. Then 
from the definition of the algorithm, the set of constraints $T$ from which
the model is obtained contains necessarily a set of constraints
constraint $T_1 \in \Tscr_1$ and $T_2 \in \Tscr_2$. Moreover, $T$
contains a set of \tsl{union} constraints specifying the split of linear
formulas among the branches.
Hence the \In{\cdot}{\cdot} elements of the model $M$ of $T$ specifies
how the formulas in the linear contexts of $\Sscr$ are to be split among
the linear contexts of $\Sscr_1$ and $\Sscr_2$. We use the same
split of formulas in the \sellf\ derivation. Finally, since the names used 
for the (linear) generic contexts appearing in one premise are different
from the names used in the other premise, 
the model $M$ restricted to the names appearing in the resulting
root side-formulas, $\Sscr_1$ and $\Sscr_2$, will also be a model of 
$T_1$ and $T_2$. Hence by induction hypothesis these are also valid
derivations in \sellf. 

Notice that the unbounded generic contexts do not 
cause a problem, although, their names are shared among the premises, since
they are subject to no constraints, except \tsl{mctx}, and that unbounded
contexts allow weakening.
\end{proof}

Completeness follows in a similar fashion, but by induction on the height
of derivations. It follows very closely the proof of the soundness lemma.
Notice that in the base case our algorithm a new set of constraints where
$F$ is any one of the subexponential 
contexts.

\begin{lemma}
Let $\Sscr$ be a root sequent context and $F$ a formula. Let $\Xi$ be a
macro-rule obtained in \sellf\ whose end sequent is $\vdash \Sscr
\Downarrow F$. Let $\mr{F}{\Cscr}{\Oscr}{\Tscr}{\Sscr}$ be the macro-rule
constructed using the algorithm above. Then there is a model $M$ of a 
set of constraints in $\Tscr$ that when used to rewrite $\Oscr$ the
resulting open premises are equivalent to those in $\Xi$. 
\end{lemma}

\end{comment}

\subsection{Putting two derivations together}

%Given a common root sequent context $\Sscr$, we will be interested in 
%determining whether if there is a proof where one formula in the
%meta-logic is focused 

Let $\mr{F_1}{\Sscr}{\Tscr_0}$ be the macro-rule of a formula $F_1$. After
finishing all possible applications of the rules, we obtain a final tuple
$\langle \Cscr_f, \Oscr_f, \Tscr_f \rangle$. Each sequent in $\Oscr_f$ is described only
by a set of generic contexts (since they are of the form $\genK : \Gamma^i
\Uparrow \cdot$), which are related through the constraints determined in
$\Tscr_f$.

Given this generic context and constraints, we can use them as an initial state
for the construction of another macro-rule of, say, a formula $F_2$:
$\mr{F_2}{\Sscr_f}{\Tscr_f}$. By the soundness and completeness of the
macro-rules, this is equivalent to a derivation in \sellf\ that focus on $F_1$
and later on $F_2$ in one of its branches.

Using this combination of macro-rules, it's possible to reason about the
permutation of inference rules of the system specified. This is described in
details in Section \ref{sec:permutations}

\section{Permutation Lemmas}
\label{sec:permutations}

\input{perm_def}

\bibliographystyle{abbrv}
\bibliography{permutation}


\end{document}
